---
title: "PA1_template.Rmd"
author: "Chris Aoyama"
output: html_document
---


##Loading and preprocessing the data
### Install and open packages
```{r echo=TRUE}
library(ggplot2)
library(knitr)
library(lubridate)
library(dplyr)
```
### Loading data
```{r echo=TRUE}
data <- read.csv("activity.csv", header = TRUE,
sep = ",", colClasses = c("numeric", "character", "integer"))
```
### Tidy data
```{r echo=TRUE}
data$date <- ymd(data$date)
```
##What is mean total number of steps taken per day?
```{r echo=TRUE}
steps <- data %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print
```
### Print histogram
```{r echo=TRUE}
ggplot(steps, aes(x = steps)) + 
  geom_histogram(fill = "green", binwidth = 1000) 
  labs(title = "Histogram of Steps per day", 
  x = "Steps per day", y = "Frequency")
```
### Find mean of steps
```{r echo=TRUE}
step_mean <- mean(steps$steps, na.rm = TRUE)
step_mean
```
### Find median of steps
```{r echo=TRUE}
step_median <- median(steps$steps, na.rm = TRUE)
step_median
```
## What is the average daily activity pattern?

### Find interval
```{r echo=TRUE}
steps_by_interval <- aggregate(steps ~ interval, data, mean)

plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")

max_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
max_interval
```

## Imputing missing values
### Total NA's
```{r echo=TRUE}
sum(is.na(data$steps))
```
### Fill in data
```{r echo=TRUE}
data_fill <- transform(data, steps = ifelse(is.na(data$steps), 
steps_by_interval$steps[match(data$interval, steps_by_interval$interval)], data$steps))
```

### New dataset using data_fill
```{r echo=TRUE}
  steps_fill <- data_fill %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print
```
### Plot histogram of the total number of steps taken
```{r echo=TRUE}
hist(steps_fill$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")

```
### Find mean
```{r echo=TRUE}
mean_steps_fill <- mean(steps_fill$steps, na.rm = TRUE)
mean_steps_fill
```
### Find median
```{r echo=TRUE}
median_steps_fill <- median(steps_fill$steps, na.rm = TRUE)
median_steps_fill
```
### Do these values differ from the estimates from the first part of the assignment?
```{r echo=TRUE}
mean_diff <- mean_steps_fill - step_mean
mean_diff
median_diff <- median_steps_fill - step_median
median_diff
```
## Are there differences in activity patterns between weekdays and weekends?

### Create new factors
```{r echo=TRUE}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday","Friday")
data_fill$dow = as.factor(ifelse(is.element(weekdays(as.Date(data_fill$date)),weekdays), "Weekday", "Weekend"))

steps_interval <- aggregate(steps ~ interval + dow, data_fill, mean)
```
###Plot weekday vs. weekend
```{r echo=TRUE}
library(lattice)

xyplot(steps_interval$steps ~ steps_interval$interval|steps_interval$dow, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")
```
